CHEC Study Sweat Chloride

### 2025-01-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
Update sweat_test_analysis_chec.qmd

---
title: "CHEC Study Sweat Chloride"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(ggpmisc)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

# CHCO data

```{r}
# Edith's data
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
  na.strings = ""
)
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
# Diagnosis per Angela
df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Numeric variables
num_vars <- c("Amount", "Na", "K", "Cl")
num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
# Filter per Maxie's code
exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
    Modulator == "None"
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Genotype info
## By severity
df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "/")
df$geno_risk <- factor(df$geno_risk)
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
df$geno_class <- factor(paste0(
  pmin(df$geno1_class, df$geno2_class), "/",
  pmax(df$geno1_class, df$geno2_class)
))
levels(df$geno_class) <- c(
  "MF/MF", "MF/MF", "MF/RF", "MF/Unknown", "MF/MF",
  "MF/RF", "MF/Unknown", "Both RF or Unknown", "Both RF or Unknown"
)
# Which arm?
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sodium to chloride ratio
df$`Na/Cl Ratio` <- df$Na / df$Cl
high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
high_cl <- unique(which(df$Cl > 160))
high_na <- unique(which(df$Na > 160))
df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
# Sweat chloride group
df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# First measure for everyone, remove those with no sweat chloride
df <- df %>%
  filter(!is.na(Cl)) %>%
  group_by(PatientID) %>%
  filter(row_number() == 1) %>%
  ungroup()
# Order groups
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF")
)
```

## By sweat chloride group

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
  geom_density(aes(group = `Sweat Chloride Group`)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
  geom_density(aes(group = `Sweat Chloride Group`)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~`Sweat Chloride Group`, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, `Sweat Chloride Group`) %>%
  tbl_summary(by = `Sweat Chloride Group`)
```

## By classification

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
  geom_density(aes(group = Diagnosis)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
  geom_density(aes(group = Diagnosis)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~Diagnosis, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, Diagnosis) %>%
  tbl_summary(by = Diagnosis)
```

## By genotype severity

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = geno_class)) +
  geom_density(aes(group = geno_class)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = geno_class)) +
  geom_density(aes(group = geno_class)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~geno_class, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, geno_class) %>%
  tbl_summary(by = geno_class)
```

# CHEC data

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
# Chloride groups
chec$swc_group <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
chec$swc_group_pm <- cut(chec$swhcavgpm, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
```

# Pre-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
  geom_density(aes(group = swc_group_pm)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
  geom_density(aes(group = swc_group_pm)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = swc_group_pm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group_pm, scales = "free", ncol = 2)
chec %>%
  select(swhsavgpm, swhcavgpm, swc_group_pm) %>%
  tbl_summary(by = swc_group_pm)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
chec %>%
  select(swhsavgpm, swhcavgpm, genotype_group) %>%
  tbl_summary(by = genotype_group)
```

# Post-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group)) +
  geom_density(aes(group = swc_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group)) +
  geom_density(aes(group = swc_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, swc_group) %>%
  tbl_summary(by = swc_group)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, genotype_group) %>%
  tbl_summary(by = genotype_group)
```

## By modulator type

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = modqual)) +
  geom_density(aes(group = modqual)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = modqual)) +
  geom_density(aes(group = modqual)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = modqual)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~modqual, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, modqual) %>%
  tbl_summary(by = modqual)
```

>>>>

---
title: "CHEC Study Sweat Chloride"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 9
    fig-height: 7
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/GitHub/styles/american-medical-association.csl
editor: source
execute:
  warning: false
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(readxl)
library(arsenal)
library(ggpubr)
library(performance)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
```

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv",
  na.strings = c("", "NA")
)
# Calculate ratios
chec$pre_ratio <- chec$swhsavgpm / chec$swhcavgpm
chec$post_ratio <- chec$swsavg / chec$swcavg
# Long dataset for easier plotting and paired analysis
chec_long <- chec %>%
  rename(
    pre_cl = swhcavgpm, pre_na = swhsavgpm, post_cl = swcavg, post_na = swsavg
  ) %>%
  pivot_longer(c(pre_cl, pre_na, post_cl, post_na, pre_ratio, post_ratio),
    names_to = c("pre_post", ".value"),
    names_pattern = "(.*)_(.*)",
    values_to = "count"
  )
chec_long$pre_post <- factor(chec_long$pre_post, levels = c("pre", "post"))
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss2", "meansd", "range"),
    cat.stats = c("N", "Nmiss2", "countpct")
  )
# Labels
label(chec_long$cl) <- "Sweat Cl"
label(chec_long$na) <- "Sweat Na"
label(chec_long$ratio) <- "Sweat Na:Cl"
```

# Tables

## Table 1: Participant demographics by CFTR modulator

```{r results='asis'}
f <- modqual~sex + age + agecat_txt + race_txt + genotype_group
t1 <- tableby(f, data = chec, control = mycontrols)
summary(t1, pfootnote = T)
```

- I think that we currently don't have ethnicity data.

## Table 2: Change in sweat electrolytes

```{r results='asis'}
pt1 <- paired(
  pre_post ~
    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
  data = chec_long, id = subject,
  control = paired.control(signed.rank.exact = F)
)
summary(pt1, pfootnote = T)
```

## Table 3: Change in sweat electrolytes by modulator

```{r results='asis'}
pt2 <- paired(
  pre_post ~
    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
  data = chec_long, id = subject, strata = modqual,
  control = paired.control(signed.rank.exact = F)
)
summary(pt2, pfootnote = T)
```

# Histograms and density plots

## Sweat chloride (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat sodium (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat sodium:chloride ratio (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat chloride by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

## Sweat sodium by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

## Sweat sodium:chloride ratio by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

# Boxplots

```{r}
ggplot(chec_long, aes(y = ratio, x = pre_post, fill = pre_post)) +
  geom_boxplot() +
  geom_signif(
    comparison = list(c("pre", "post")), test = "wilcox.test",
    test.args = list(paired = TRUE, exact = FALSE), y_position = 3.5,
    map_signif_level = function(p) format.pval(p, digits = 3, eps = 0.001)
  ) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

# Scatterplots and models

To evaluate whether the relationship between pre-modulator sweat chloride and pre-modulator sodium:chloride ratio is logarithmic, we compared three models:

1. A simple linear model with no transformations.
2. A linear model with log-transformed sweat chloride as the outcome.
3. A quadratic model.

The model residuals do *not* appear to be improved by log transforming the outcome. Also, the quadratic model is better than the linear model by BIC, but this does not remain the case when the influential point (CFF ID 1650133) is removed.

```{r}
pre_model_df <- chec %>%
  select(cffid_first, swhcavgpm, pre_ratio) %>%
  drop_na() %>%
  unique()

m1 <- lm(swhcavgpm ~ pre_ratio, data = pre_model_df)
log_m <- lm(log(swhcavgpm) ~ pre_ratio, data = pre_model_df)
m2 <- lm(swhcavgpm ~ pre_ratio + I(pre_ratio^2), data = pre_model_df)
```

## Linear model assumptions check

```{r}
check_model(m1)
```

## Log model assumptions check

```{r}
check_model(log_m)
```

## Quadratic model assumptions check

```{r}
check_model(m2)
```

## Model predicted values

```{r}
ggplot(pre_model_df, aes(y = swhcavgpm, x = pre_ratio)) +
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(color = "Linear")) +
  geom_smooth(
    method = "lm", formula = y ~ x + I(x^2), se = F,
    aes(color = "Quadratic")
  ) +
  geom_smooth(
    method = "glm", formula = y ~ x, se = F, aes(color = "Log Linear"),
    method.args = list(family = gaussian(link = "log"))
  ) +
  theme_bw() +
  theme(legend.title = element_blank())
```
